---
title: "Homework 2"
author: Safiya Sirota
date: 2021-10-09
output: 
  github_document:
    toc: TRUE
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

## Problem 1

__Importing Mr. Trash Wheel dataset:__

```{r}
mrtrash_df <- 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N406"
    ) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls,0))
```

__Importing precipitation datasets:__

2018 dataset:

```{r}
prcp_2018 <-
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month, total) %>% 
  mutate(year = 2018)
```

2019 dataset:

```{r}
prcp_2019 <-
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2019 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month, total) %>% 
  mutate(year = 2019)
```

Combining and tidying the two prcp datasets:

```{r}
prcp_tidy <-
  bind_rows(prcp_2018, prcp_2019) %>% 
  mutate(month = month.name[month]) %>% 
  select(year, month, prcp_amt = total)
```

__Describing the datasets:__


There are `r nrow(prcp_tidy)` observations in the `prcp_tidy` dataset and `r nrow(mrtrash_df)` observations in the `mrtrash_df` dataset. The key variables in the `mrtrash_df` dataset are:

* __dumpster identification__ (`r names(mrtrash_df)[1:4]`),

* __amount of trash__ (`r names(mrtrash_df)[5:6]`),

* __types of trash__ (`r names(mrtrash_df)[7:13]`), and

* __number of homes powered__ (`r names(mrtrash_df)[14]`)

The key variables in the `prcp_tidy` dataset are:

* __`r names(prcp_tidy)[1:2]`__ and

* __precipitation amount__.

For available data, the total precipitation in 2018 was `r sum(pull(prcp_2018, total))` inches. The median number of sports balls in a dumpster in 2019 was `r median(pull(mrtrash_df, sports_balls))`.
