---
title: "Homework 2"
author: Safiya Sirota
date: 2021-10-09
output: 
  github_document:
    toc: TRUE
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)
```

## Problem 1

__Importing Mr. Trash Wheel dataset:__

```{r}
mrtrash_df <- 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N406"
    ) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls,0))
```

__Importing precipitation datasets:__

2018 dataset:

```{r}
prcp_2018 <-
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month, total) %>% 
  mutate(year = 2018)
```

2019 dataset:

```{r}
prcp_2019 <-
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
    sheet = "2019 Precipitation",
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month, total) %>% 
  mutate(year = 2019)
```

Combining and tidying the two prcp datasets:

```{r}
prcp_tidy <-
  bind_rows(prcp_2018, prcp_2019) %>% 
  mutate(month = month.name[month]) %>% 
  select(year, month, prcp_amt = total)
```

__Describing the datasets:__


There are `r nrow(prcp_tidy)` observations in the `prcp_tidy` dataset and `r nrow(mrtrash_df)` observations in the `mrtrash_df` dataset. The key variables in the `mrtrash_df` dataset are:

* __dumpster identification__ (`r names(mrtrash_df)[1:4]`),

* __amount of trash__ (`r names(mrtrash_df)[5:6]`),

* __types of trash__ (`r names(mrtrash_df)[7:13]`), and

* __number of homes powered__ (`r names(mrtrash_df)[14]`)

The key variables in the `prcp_tidy` dataset are:

* __`r names(prcp_tidy)[1:2]`__ and

* __precipitation amount__.

For available data, the total precipitation in 2018 was `r sum(pull(prcp_2018, total))` inches. The median number of sports balls in a dumpster in 2019 was `r median(pull(mrtrash_df, sports_balls))`.

## Problem 2

__Importing and cleaning `pols-month` dataset:__

```{r}
pols_df <- read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    year = as.numeric(year), 
    month = month.name[as.numeric(month)],
    day = as.numeric(day),
    president = case_when(
      prez_gop == 1 ~ "gop",
      prez_dem == 1 ~ "dem")
    ) %>% 
  select(-prez_gop, -prez_dem, -day)
```

__Importing and cleaning `snp` dataset:__

```{r}
snp_df <- read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  mutate(
    year = case_when(
      year < 22 ~ 2000 + as.numeric(year), 
      year > 22 ~ 1900 + as.numeric(year)
    ),
    month = month.name[as.numeric(month)],
    day = as.numeric(day)
  ) %>% 
  select(year, month, close)
```

__Importing and cleaning `unemployment` dataset:__

```{r}
unemploy_df <- read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemp_pct"
  ) %>% 
  mutate(month = month.name[match(month, month.abb)]) %>% 
  rename(year = Year)
```

__Joining all three datesets:___
```{r}
five38_df <-
  full_join(pols_df, snp_df, by = c("year", "month")) %>% 
  full_join(unemploy_df,by = c("year", "month"))
```

__Desribing the datasets:__


## Problem 3

__Importing and tidying `popular_baby_names` dataset:__

```{r}
names_df <- read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    gender = str_to_title(gender),
    ethnicity = str_to_title(ethnicity),
    childs_first_name = str_to_title(childs_first_name),
    ethnicity = recode(ethnicity, 
                       "Asian And Paci" = "Asian And Pacific Islander",
                       "Black Non Hisp" = "Black Non Hispanic",
                       "White Non Hisp" = "White Non Hispanic")
  ) %>% 
  distinct(year_of_birth, rank, ethnicity, childs_first_name, .keep_all = TRUE)
```

The case structure of `childs_first_name` was inconsistent throughout the data, but we were able to fix this with the `str_to_title` function.  

We could also see that the values of `ethnicity` were not consistent throughout the dataset, with API, non-hispanic black, and non-hispanic white indicators changing throughout. We were able to fix this by recoding the ethnicity variable values.  

Lastly, we needed to get rid of duplicates. The `distinct()` function is useful for this.

__Creating a readable `olivia` table__:

```{r}
olivia <- 
  filter(names_df, 
         childs_first_name == "Olivia" 
         & gender == "Female") %>% 
  select(ethnicity, rank, year_of_birth) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  )
```

__Creating a readable table for the most popular male name__:

First, we should determine the most popular male name over time. I will define this as the male name with highest count across all years in the dataset. We can make a table to help:

```{r}
pop_name <-
  filter(names_df, gender == "Male") %>% 
  select(childs_first_name, count) %>% 
  group_by(childs_first_name) %>% 
  summarize_all(sum) %>% 
  arrange(desc(count))
```

We find that the most popular male name is Ethan, so now we will make an `ethan` table.

```{r}
ethan <- 
  filter(names_df, 
         childs_first_name == "Ethan" 
         & gender == "Male") %>% 
  select(ethnicity, rank, year_of_birth) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  )
```

__Making a scatterplot of `rank` vs. `count`:__

```{r}
names_df %>% 
  filter(gender == "Male" & ethnicity == "White Non Hispanic") %>% 
  ggplot(aes(x = rank, y = count)) +
  geom_point()
```

